---
title: "INFS 691 - Homework8 - Part1"
subtitle: Text Mining and Sentiment Analysis
author: "Rose Dunderdale"
output:
  html_document: default
  html_notebook: default
---

Read the document "Homework8 Instructions.docx" for detailed 
information about Homework 8 that has three parts.

In "Homework8.Part1", you will carry out text mining and sentiment 
analysis for the text data "GuderLUC.txt". This text data contains
the actual teacher/course evaluations for the courses that I taught 
at Loyola University of Chicago (LUC) in the last 3 years. You need
to perform all steps of text mining as was done in the example problem 
presented in the reading document and in the videos.
```{r}
install.packages("tm")        
install.packages("dplyr")     
install.packages("stringr")   

install.packages("wordcloud")  
install.packages("wordcloud2") 
install.packages("syuzhet")   
install.packages("tidytext")   

library(tm)  # major package for text mining           
library(dplyr) # data manipulation          
library(stringr) # for string-split          
library(wordcloud)  # wordcloud generator          
library(wordcloud2) # wordcloud2 generator    
library(syuzhet) # sentiment lexicons             
library(tidytext) # manipulation, sentiment            
```

### Text Mining and Sentiment Analysis Process
#### Step 1. Read Data and Create Corpus 
#### Step 2. Clean Data 
#### Step 3. Create Term Document Matrix (tdm)
#### Step 4. Analyze Data
####        (word frequency and Sentiment Analysis)


#### Step 1. Read Text Data and Create Corpus
(a) Read the text data 
Text data "GuderLUC691.txt" contains the actual teacher/course evaluations for the course (INFS691) that I taught 
at Loyola University of Chicago (LUC) in the last 3 years. 
```{r}
text0 <- readLines("GuderLUC691.txt")
text0
# text0: A character vector with 196 elements.

# convert text0 to a character vector of one element (combine reviews)
text <- paste(text0, collapse=" ") # combine all in one text
text
```

(b) Create corpus
corpus is a data structure (list) that is used by the "tm" package to process and analyze text data
```{r}
corpus <- Corpus(VectorSource(text)) # corpus is a list of 2 elements 
inspect(corpus) # or corpus$content to display the content of corpus

#result: the corpus has the same content as the text vector
```

#### Step 2. Clean Data (corpus)
Purpose: to understand and analyze opinions and feelings of the customers and take actions accordingly. 
Cleaning: removing data/words that do not reflect any opinions or feelings. 

See the corpus (before cleaning)
```{r}
inspect(corpus)
```

To clean data in corpus, we will use tm_map() function
available in "tm" package
```{r}
# convert the text to lower case 
corpus <- tm_map(corpus, tolower)
inspect(corpus) # to display the corpus 

# remove english common stopwords (is, are, this, but, and, etc.)
corpus <- tm_map(corpus, removeWords, stopwords("english"))
inspect(corpus) # to display the corpus 

# remove punctuation
corpus <- tm_map(corpus, removePunctuation) #removes punctuation
inspect(corpus) # to display the corpus 

# remove additional words 
#specified words: professor , guder, course, class
corpus <- tm_map(corpus, removeWords, c("professor", "guder", "course", "class"))
inspect(corpus) # to display the corpus

# remove numbers 
corpus <- tm_map(corpus, removeNumbers)
inspect(corpus) 

# Eliminate Whitespace
corpus <- tm_map(corpus, stripWhitespace)
inspect(corpus) # to display the corpus
```

#### Create TWO objects: "text2" and "words"
    (These objects will be used in Step 4 - Text analysis)

First object: text2
(1) text2: split the clean text (corpus) into individual elements (words) and assigns them to the vector "text2".
    "Bag of words". 
```{r}
text2 <- str_split(corpus$content, pattern ="\\s+") # text2 is a list
text2 <- unlist(text2) #text2 is now a vector
#inspect(corpus)
text2
#Display text2
```

Second Object: words 
(2) words: the list of the words in text2 vector is
      is expressed in a data frame structure. 
```{r}
words <- data.frame(word=text2)
words
```

#### Step 3. Build a TermDocumentMatrix (tdm)
Text mining processes transforms unstructured text into a structured format (into a table/matrix) 
to identify meaningful patterns and new insights. 

tdm turns the corpus into a document term matrix (structured data) 
Create tdm and the following objects:
tdm_m, tdm_s, and tdm_df
```{r}
tdm <-TermDocumentMatrix(corpus) # tdm is a list 
tdm_m <- as.matrix(tdm) # change td to matrix
tdm_m
# order the tdm_m matrix by decreasing value of frequencies
# and assign them to the vector "tdm_s"
tdm_s <- sort(rowSums(tdm_m), decreasing=TRUE)
tdm_s
# tdm_s is a vector 
# convert the tdm_s vector into a data frame format (tdm_df)
# with the column names as follows:
# column1 is named as 'word', column2 is named as 'freq'
# This is needed only if you wanted to display wordcloud2 in step 4
tdm_df <- data.frame(word = names(tdm_s), freq=tdm_s)
tdm_df
```

#### Step 4. Analyze Text Data 
(1) Visual Display of word frequencies using wordcloud 
(2) Visual Display of word frequencies using barplot
(3) Sentiment Analysis - calculate sentiments (customer opinions)

#### (1) Display word frequencies using wordcloud
```{r}
# wordcloud (corpus)
set.seed(123)        # to obtain the same display      
wordcloud(corpus,    #clean corpus      
         max.words=150,  # max number of words   
         min.freq=1,     # words with the frequency below will not be plotted 
         random.order=F, # the words displayed in random order  
         color=rainbow(3),  # rainbow color option
         scale=c(4.0,0.1),  # the range of the size of the words 
         rot.per=0.30)      # proportion of the words to be rotated
```

Another package - wordcloud2
Using wordcloud2
```{r}
set.seed(123)
wordcloud2(tdm_df,
           size=0.6,
           color=rainbow(5),
           shape="oval",
           rotateRatio=0.3)
```
#### (2) Display word frequencies using a barplot 
Display barplot for the top 15 most frequent words 
```{r}
# using the tdm_s: sorted tdms_s matrix 
barplot(head(tdm_s,15),    # the object should be a (sorted) vector
        las=2,
        col=rainbow(50),
        ylab="Frequency",
        main= "Top 15 most frequently used words")
```
#### (3) Sentiment Analysis 
# sentiment analysis = the process of determining whether a text is posituve, 
# negative or neutral
#### Sentiment Scores using "bing" lexicon, which categorizes words into 
# positive and negative categories
#### Read the lexicon by bing: list of positive and negative terms 

Pull out only sentiment words.
#### Read the sentiment words in the "bing" lexicon:
```{r}
bing <- get_sentiments(lexicon="bing") # from tidytext package 
bing                                   # list of word in bing lexicon - 6786 words
words                                  # list of (bag) words in clean opus - 
                                       ## created in step 2
```

#### Determine the sentiment words in our text (corpus)
```{r}
text_bing <- inner_join(words, bing, by="word")
text_bing   # list of all sentiment words in clean corpus
```

Separate the list of the positive an negative words
```{r}
poswords <- subset(text_bing, sentiment=="positive")
negwords <- subset(text_bing, sentiment=="negative")
poswords    # display the positive words
negwords    # display the negative words 
```

Count the number of positive and negative
and calculate the sentiment score for the text.
```{r}
countPos <- nrow(poswords)    # count the number of positive words 
countNeg <- nrow(negwords)    # count the number of negative words 
countPos 
countNeg 

# Calculate sentiment score (sentiment score = countPos - countNeg)
result <- c(positive=countPos, negative=countNeg, sentimentsScore=countPos - countNeg)
result
```
This sentiment score indicates a very strong positive review by the students

### Note about the sentiment scores
The # of positive and negative sentiments, and the sentiment score depends on the number of reviews. 

The # of the positive and negative words tend to increase together for a longer review. Therefore, it might be helpful to use the ratios of positive and negative word counts if you are comparing the sentiment scores for 2 different products or services with different number of reviews (texts), it might be helpful to use the ratios of the number of positive and negative words. 

### results review
Degree of positivity/negativity = countPos/(countPos + countNeg)

Degree of positive = 398/453 = 87.86%, meaning that 87.86% of the sentiment words
in the reviews are positive. This is a high percenttage of positive words. 

OR 

Degree of positivity = countPos/countNeg

Thsi ratio is an example equal to 398/55 = 7.24, meaning that the reviews in this example included 7.24
positive words for each negative word. Again, this ratio indicates a very strong positive sentiment. 




#### Create the html file and submit it on Sakai

You need to create the html file: After completing your Rmd file,
you will execute all the chunks, save the file, and generate the 
html file containing the codes, outputs, and the answers. 

To create the html file, complete the following steps:

Step 1. (After writing all the codes) Execute all chunks one by one
starting from the first chunk (regardless of  whether you executed 
the codes while writing your codes).

Step 2. Save your file->Save As…

If you have completed Step 1, then Step 2 will save your Rmd file, 
and (at the same time) it will create the html file on your working 
directory. You will attach and submit this html file on Sakai as your homework8-Part1.

Important: Please check the html file to make sure that it contains 
           the R codes, outputs, and your answers. If not, repeat 
           Step 1 and Step 2.



#### Create the html file and submit it on Sakai

You need to create the html file: After completing your Rmd file,
you will execute all the chunks, save the file, and generate the 
html file containing the codes, outputs, and the answers. 

To create the html file, complete the following steps:

Step 1. (After writing all the codes) Execute all chunks one by one
starting from the first chunk (regardless of  whether you executed 
the codes while writing your codes). 

Step 2. Save your file->Save As…

If you have completed Step 1, then Step 2 will save your Rmd file, 
and (at the same time) it will create the html file on your working 
directory. You will attach and submit this html file on Sakai as your homework8-Part1.

Important: Please check the html file to make sure that it contains 
           the R codes, outputs, and your answers. If not, repeat 
           Step 1 and Step 2.
           
